DATA 621 Week #2 Textbook Exercises

### 2019-09-07 DATA621/Week 02 Textbook Exercises.Rmd
More work on Week 2 Textbook Exercises

---
title: "DATA 621 Week #2 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---

@@

```{r}
ggplot(playbill, aes(x=LastWeek/1000, y=CurrentWeek/1000)) +
  geom_point() +
  labs(title = "Gross Box Office Results",
       subtitle = "(in thousands of dollars)",
       x = "Previous Week",
       y = "Current Week") +
  geom_abline(slope=1, intercept = 0, color = "red")
```

@@

@@

>>>>

---
title: "DATA 621 Week #2 Textbook Exercises"
author: "Mike Silva"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    code_folding: "hide"
---

@@

```{r}
ggplot(playbill, aes(x=LastWeek/1000, y=CurrentWeek/1000)) +
  geom_point() +
  labs(title = "Gross Box Office Results",
       subtitle = "(in thousands of dollars)",
       x = "Previous Week",
       y = "Current Week") +
  geom_abline(slope=1, intercept = 0, color = "red") + 
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90"))
```

@@

The dataset *prostate* comes from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy.  Fit a model with *lpsa* as the response and *lcavol* as the predictor.  Record the residual standard error and the $R^2$.  Now add *lweight*, *svi*, *lbph*, *age*, *lcp*, *pgg45* and *gleason* to the model on one at a time.  For each model record the residual standard error and the $R^2$.  Plot the trends in these two statistics.

```{r}
prostate <- read.delim("data/prostate.txt")

get_row <- function(fit, n_vars){
  r2 <- summary(fit)$r.squared
  rse <- sigma(fit)
  data.frame("Predictors" = n_vars, "RSE" = rse, "R2" = r2)
}

df <- get_row(lm(lpsa ~ lcavol, prostate), 1) %>% 
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight, prostate), 2)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi, prostate), 3)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph, prostate), 4)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph, prostate), 5)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age, prostate), 6)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp, prostate), 7)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45, prostate), 8)) %>%
  bind_rows(get_row(lm(lpsa ~ lcavol + lweight + svi + lbph + age + lcp + pgg45 + gleason, prostate), 9)) %>%
  gather(Measure, Value, -Predictors) %>%
  mutate(Measure = as.factor(Measure))

ggplot(df, aes(x=Predictors, y = Value, color = Measure)) +
  geom_line() +
  scale_x_continuous(breaks = round(seq(min(df$Predictors), max(df$Predictors), by = 1), 0)) +
  scale_y_continuous(breaks = round(seq(0, 1, by = 0.05), 2)) +
  annotate("text", x = 5, y = max(df[df$Measure == "R2",]$Value) + 0.01, label ="R2") +
  annotate("text", x = 5, y = min(df[df$Measure == "RSE",]$Value) + 0.02, label ="RSE") +
  labs(title="R2 and RSE by Number of Variables", 
       x = element_blank(), 
       y = element_blank()) +
  theme(legend.position = "None",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90"))
```

@@

Using the *prostate* data, plot *lpsa* against *lcavol*.  Fit the regressions of *lpsa* on *lcavol* and *lcavol* on *lpsa*.  Display both regression lines on the plot.  At what point do the two lines intersect? **They intersect where the cancer volume roughly 1.3 and the log of PSA is 2.5**

```{r}
fit <- lm(lpsa ~ lcavol, prostate)
df <- data.frame(lcavol = prostate$lcavol, 
                 lpsa = predict(fit, prostate), 
                 Model = "lpsa ~ lcavol")

fit <- lm(lcavol ~ lpsa, prostate)
df <- df %>% 
  bind_rows(data.frame(lcavol = predict(fit, prostate), 
                       lpsa = prostate$lpsa, 
                       Model = "lcavol ~ lpsa")) %>%
  mutate(Model = as.factor(Model))

ggplot(prostate, aes(lcavol, lpsa)) + 
  geom_point() + 
  geom_line(aes(lcavol, lpsa, color = Model), data = df) + 
  labs(x = "Cancer Volume", 
       y = "PSA (log)") +
  theme(legend.position = "bottom",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray90")) +
  scale_x_continuous(breaks = round(seq(min(prostate$lcavol), max(prostate$lcavol), by = 1), 0)) +
  scale_y_continuous(breaks = round(seq(min(prostate$lpsa), max(prostate$lpsa), by = 1), 0)) 
```
